Cultural evolutionary dynamics with prestige bias

### 2022-10-27 ms/ms.tex

\documentclass[12pt]{extarticle}
\usepackage{geometry}
\geometry{
a4paper,
total={170mm,257mm},
left=20mm,
top=20mm,
headheight=12pt
}

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
% appendix
\usepackage[title,page]{appendix}
\usepackage{chngcntr}

@@

% line numbers
\usepackage[displaymath, mathlines]{lineno}
 \renewcommand\linenumberfont{\normalfont\small\sffamily}
%\linenumbers
\modulolinenumbers[2]

% Supplementary
% https://support.authorea.com/en-us/article/how-to-create-an-appendix-section-or-supplementary-information-1g25i5a/
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
		\setcounter{equation}{0}
        \renewcommand{\theequation}{A\arabic{equation}}%
}

@@

% line numbers
\usepackage[displaymath, mathlines]{lineno}
\renewcommand\linenumberfont{\normalfont\small\sffamily}
\linenumbers
\modulolinenumbers[2]

% Title page
\title{Cultural evolutionary dynamics with prestige bias}

% Authors
\renewcommand\Affilfont{\small}

\author[1]{Saar Egozi}
\author[2,3,$\dagger$]{Yoav Ram}

\affil[1]{School of Computer Science, Reichman University, Herzliya, Israel}
\affil[3]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv, Israel}
\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv, Israel}
%\affil[*]{These authors contributed equally to this work}
\affil[$\dagger$]{Corresponding author: yoav@yoavram.com}

% Document
\begin{document}
\maketitle

\begin{abstract}
Copying role-models can be an efficient method for acquiring knowledge. A common bias when choosing a role-model to copy is success bias: copying whoever appears more successful. This bias depends on the performance of the role-model alone, with no other factors. We propose an additional bias that may be prevalent in cultural transmission: influence bias, in which role-model choice is affected by the number of individuals that have already copied each potential role-model. We combine success and influence bias together to a ``prestige bias'' and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations. We find analytic approximations to our stochastic model, facilitating further mathematical analysis and reducing the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading favored (by bias) cultural trait, in both constant and changing environments, which resemble Kimura's approximations for population-genetic models.
These approximations show that success bias effectively plays the part of natural selection, whereas influence bias effectively reduces the population size.	
Influence bias also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a good description of cultural transmission, 
especially in human societies where social media is popular. 
Further work is needed to test if this model could predict various phenomena in cultural evolution when extended with the effects of selection and innovation.
\end{abstract}

@@

%% Introduction
\section*{Introduction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Models}

@@

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by the estimated biased trait value $\beta(A_{i,j})$ and the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} of \citet{evolutionBook} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}
Here, the bias weight $\alpha_{ij}$ is a characteristic of the interaction of role-model~$j$ with copier $i$, that determines the relative significance of success and influence within the role-model's overall prestige, in the eyes if the copier. The trait of role-model $j$ estimated by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeTrait}, and $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$ for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + X\alpha'\beta(\hat{A})} = \frac{(N-X)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X}
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.

Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only (contrary to $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.
\paragraph{Generalized binomial approximation.} 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
(i) the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$); and
(ii) the expected number of copiers of each role-model equals its relative biased trait value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated trait value.

\paragraph{Dirichlet-Multinomial approximation.}
The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$), meaning there's no meaning for the order of copiers. 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the trait values $A_j$ and the bias weights $\alpha_j$, as the estimation error is uniform, meaning all copiers are identical.

%%%%%%%%%%%%%
\subsection*{Generalized binomial approximation}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

\begin{result}[Generalized binomial approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

@@

This result gives the following two corollaries on the expected number of followers of a given role-model by the end of the role-model choice process. The proof of \cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

@@

\begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{../figures/final/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation of the GB approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

Although basic, \cref{fig:GBD} shows good fit of the GB approximation.
This validation is initial, and the more extensive validations we do on the Dirichlet-Multinomial approximation, because it is what we will use in our analysts.

%%%%%%%%%%%%%
\subsection*{Dirichlet-Multinomial distribution approximation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}

After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and fixation time of a favored phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and influence on prestige bias, given by \cref{eq:prestige}.
For simplicity, we do not include role-model estimation error in this analysis, i.e $e_i=0$ for every copier $i$.
As shown above, transmission in our model is approximately DM distributed (\cref{cor:dirichlet} and \cref{eq:copier_choose}).

We start by finding the expectation and variance of the change in frequency from one generation to the next, which are the drift and diffusion terms of the diffusion equation.
The proof is in Appendix~\ref{sec:drift_diff_const}.
\\

@@

@@

Using our effective selection coefficient, $1-\beta$, and effective population size, $N_e$, with the population-genetics fixation probability approximation given by \citet[eq.~8]{kimura}, we get the following result.\\

\begin{result}[Fixation probability]
The fixation probability is approximately
\begin{equation}\label{eq:kimura_p}
\pi = \frac{1-e^{-2(1-\beta)N_e x}}{1-e^{-2(1-\beta)N_e}}
\end{equation}
where $x$ is the initial frequency of the favored phenotype $\hat{A}$.
\end{result}

@@

\paragraph*{Changing environment}. After finding a good approximation in constant environment, where the favorable trait is always $\hat{A}$, we proceeded to find an approximation for a changing environment. 
Following \citet{changeEnv}, we find the following expressions for the expectation and variance of the change in frequency between $t$ generations. 
The proof is in Appendix~\ref{sec:drift_diff_chang}.
\\

@@

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process, demonstrating that influence becomes more favored as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be copied more often due to transmission biases. 
A common bias is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested that because it is hard to estimate success, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful in other traits \citep{fijian_social_bias}.

@@

Prestige bias can help to cheaply estimate and acquire knowledge, which may facilitate survival and reproduction. However, it is not always the case, and there could be negative repercussions to this bias, such as invasion of maladaptive traits.
\citet{best_of_k} mention that social learning not only takes the form of random copying of other individuals, but also involves learners' choice of what to learn and from whom to learn. They suggest a best-of-k model where an individual samples~$k$ role-models and chooses the one he deems most "successful". They mentioned that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They developed new mathematical models that are simpler and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff. Our model, which includes influence bias, is consistent with this claim. When a maladaptive trait is "piggybacking" a role-model with high influence, this trait could spread in the population.
In addition, they show that best-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning. 
Our model only includes social learning, and not individual learning, but it could be interesting to combine it with individual learning and see how it affects the dynamics.

@@

{\small
\section*{Acknowledgements}
We thank Marcus Feldman, Martin Pontz, and Tal Simon for discussions and comments.
This work was supported in part by 
the Israel Science Foundation (YR 552/19),
Minerva Stiftung Center for Lab Evolution (YR),
and 
the John Templeton Foundation (YR 61809).
}

@@

\section*{References}
\nolinenumbers
\bibliography{refs}
%\bibliographystyle{agsm}
\bibliographystyle{apalike}

\begin{appendices}
\renewcommand{\theequation}{\thesection\arabic{equation}}
\counterwithin*{equation}{section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General binomial distribution approximation} \label{sec:GBD}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drift and diffusion in a constant environment} \label{sec:drift_diff_const}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drift and diffusion in a changing environment} \label{sec:drift_diff_chang}
\paragraph{Proving drift and diffusion terms in a changing environment.}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation*}
E[X_t/N - x] \simeq x(1-x) S_t / N_e \;, 
\quad
\text{and}
\quad
V(X_t/N-x) \simeq  t x(1-x) / N_e \;,
\end{equation*}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.

@@

\end{appendices}

%%%
\end{document}

>>>>

\documentclass[12pt]{extarticle}

\makeatletter

\def\blfootnote{\gdef\@thefnmark{}\@footnotetext}

\makeatother

\usepackage{geometry}
\geometry{
a4paper,
total={170mm,257mm},
left=20mm,
top=20mm,
headheight=12pt
}

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage{cjhebrew}

@@

% line numbers
\usepackage[displaymath, mathlines]{lineno}
 \renewcommand\linenumberfont{\normalfont\small\sffamily}
%%%%\linenumbers
\modulolinenumbers[2]

% Supplementary
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

\begin{document}

\begin{titlepage}
	\centering
	\includegraphics[width=0.4\textwidth]{ReichmanLogo}\par\vspace{1cm}
	{\huge Reichman University \par}
	{\Large Efi Arazi School of Computer Science \par}
	{\Large M.Sc. Program - Research Track \par}

	\vspace{1.5cm}
	{\Huge Cultural Evolutionary Dynamics with Success and Influence Bias\par}
	\vspace{1.5cm}
	{\large by\par}
	{\Large\bfseries Saar Egozi\par}
	\vspace{1cm}
	{\large Supervised by Dr. Yoav Ram\par}
	{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv 6997801, Israel\par}
%	{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv 6997801, Israel \par}

	\vspace{1cm}
	{M.Sc. dissertation, submitted in partial fulfillment of the requirements\par}
	{for the M.Sc. degree, research track, School of Computer Science\par}
	{Reichman University}

	\vspace{1cm}
	% Bottom of the page
	{\large \today \par}

\end{titlepage}

% title page stuff
%\title{Cultural Evolutionary Dynamics with Success and Influence biases}

%\renewcommand\Affilfont{\small}
%\author[1]{Saar Egozi}
%\author[2,3,*]{Yoav Ram}
%\affil[1]{School of Computer Science, Reichman University, Herzliya 4610101, Israel}
%\affil[2]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv 6997801, Israel}
%\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv 6997801, Israel}
%\affil[*]{Corresponding author: yoav@yoavram.com, ORCID 0000-0002-9653-4458}

%\date{\today}

%\maketitle

This work was carried out under the supervision of Dr. Yoav Ram as part of the M.Sc.
program of Efi Arazi School of Computer Science, Reichman University.

@@

\section{Abstract}
Copying role-models can be an efficient method for acquiring knowledge. A common bias when choosing a role-model to copy is success bias: copying whoever appears more successful. This bias depends on the performance of the role-model alone, with no other factors. We propose an additional bias that may be prevalent in cultural transmission: influence bias, in which role-model choice is affected by the number of individuals that have already copied each potential role-model. We combine success and influence bias together to a ``prestige bias'' and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations. We find analytic approximations to our stochastic model, facilitating further mathematical analysis and reducing the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading favored (by bias) cultural trait, in both constant and changing environments, which resemble Kimura's approximations for population-genetic models.
These approximations show that success bias effectively plays the part of natural selection, whereas influence bias effectively reduces the population size.	
Influence bias also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a good description of cultural transmission, 
especially in human societies where social media is popular. 
Further work is needed to test if this model could predict various phenomena in cultural evolution when extended with the effects of selection and innovation.

\blfootnote{This work is in preparation for submission to a high-impact journal.}
\pagebreak

\tableofcontents

\section{Introduction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Models and methods}

@@

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by the estimated biased trait value $\beta(A_{i,j})$ and the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} of \citet{evolutionBook} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}

Here, the bias weight $\alpha_{ij}$ is a characteristic of the interaction of role-model~$j$ with copier $i$ that determines the relative significance of success and influence in the role-model's overall prestige in the eyes of the copier.

We assume every individual evaluates the importance of success and influence differently. Additionally, we assume every role-model displays his influence and success individually, for example - a person with more followers but lack of actual skill may emphasize the number of his followers rather than his actual skill. We therefore modeled $\alpha_{ij}$ to be the relation between every copiers intrinsic significance of estimation, and every role-models personal emphasis.

The trait of role-model $j$ estimated by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeTrait}, and $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.

\subsection{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$ for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + C\alpha'\beta(\hat{A})} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + \alpha'C}
\end{equation}
where $C$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\subsection{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only (contrary to $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$). This means that there's no actual estimation error in the population, just a fixed bias in every copier estimation.
\item the expected number of copiers of each role-model equals its relative biased trait value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated trait value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$), meaning there's no meaning for the order of copiers. 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the trait values $A_j$ and the bias weights $\alpha_j$, as the estimation error is uniform, meaning all copiers are identical.
\end{enumerate}

\subsection{Generalized binomial distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

\begin{result}[GBD approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

@@

@@

\begin{proof}\cref{append:GBD}
\end{proof}

@@

\begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{../figures/final/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation of the GBD approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

Although basic, \cref{fig:GBD} shows good fit of the GBD approximation.
This validation is initial, and the more extensive validations we do on the Dirichlet-Multinomial approximation, because it is what we will use in our analysts.

\subsection{Dirichlet-Multinomial distribution}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fixation probability and time}

After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditioned fixation time (conditioned on the population reaching fixation) of a favored phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and influence on prestige bias, given by \cref{eq:prestige}.
For simplicity, we do not include role-model estimation error in this analysis, i.e $e_i=0$ for every copier $i$.
As shown above, transmission in our model is approximately DM distributed (\cref{cor:dirichlet} and \cref{eq:copier_choose}).

We start by finding the expectation and variance of the change in frequency from one generation to the next, which are the drift and diffusion terms of the diffusion equation.\\

@@

\begin{proof}
\cref{append:drift_diff_const}
\end{proof}

@@

Note that when $N>>1$, $N_e \approx \alpha N$, resulting in a very convenient expression.\\

Using our effective selection coefficient, $1-\beta$, and effective population size, $N_e$, with the population-genetics fixation probability approximation given by \citet[eq.~8]{kimura}, we get the following result:
\begin{result}[Fixation probability]
The fixation probability is approximately
\begin{equation}\label{eq:kimura_p}
\pi = \frac{1-e^{-2(1-\beta)N_e x}}{1-e^{-2(1-\beta)N_e}}
\end{equation}
where $x$ is the initial frequency of the favored phenotype $\hat{A}$.
\end{result}

@@

\paragraph*{Changing environment}. After finding a good approximation in constant environment, where the favorable trait is always $\hat{A}$, we proceeded to find an approximation for a changing environment. 
Following \citet{changeEnv}, we find the following expressions for the expectation and variance of the change in frequency between $t$ generations.\\

@@

\begin{proof}
\cref{append:drift_diff_chang}
\end{proof}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/final/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process, demonstrating that influence becomes more favored as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Discussion}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be copied more often due to transmission biases. 
A common bias is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested that because it is hard to estimate success, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful in other traits \citep{fijian_social_bias}.

@@

Prestige bias can help to cheaply estimate and acquire knowledge, which may facilitate survival and reproduction. However, it is not always the case, and there could be negative repercussions to this bias, such as invasion of maladaptive traits.
\citet{best_of_k} mention that social learning not only takes the form of random copying of other individuals, but also involves learners' choice of what to learn and from whom to learn. They suggest a best-of-k model where an individual samples $k$ role-models and chooses the one he deems most "successful". They mentioned that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They developed new mathematical models that are simpler and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff. Our model, which includes influence bias, is consistent with this claim. When a maladaptive trait is "piggybacking" a role-model with high influence, this trait could spread in the population.
In addition, they show that best-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning. 
Our model only includes social learning, and not individual learning, but it could be interesting to combine it with individual learning and see how it affects the dynamics.

@@

{\small
\section{Acknowledgements}
We thank Marcus Feldman, Martin Pontz, and Tal Simon for discussions and comments.
This work was supported in part by 
the Israel Science Foundation (YR 552/19),
Minerva Stiftung Center for Lab Evolution (YR),
and 
the John Templeton Foundation (YR 61809).
}

\section{References}
\nolinenumbers
\bibliography{refs}
%\bibliographystyle{agsm}
\bibliographystyle{apalike}

@@

\appendix
\section{General Binomial Distribution approximation proof}\label{append:GBD}

@@

\section{Drift and diffusion in a constant environment proof}\label{append:drift_diff_const}

@@

\section{Drift and diffusion in a changing environment proof}\label{append:drift_diff_chang}
\paragraph{Proving drift and diffusion terms in a changing environment.}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation*}
E[X_t/N - x] \simeq x(1-x) S_t / N_e \;, 
\quad
\text{and}
\quad
V(X_t/N-x) \simeq  t x(1-x) / N_e \;,
\end{equation*}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.

@@

\end{document}
